Targeted interrogation of mutational hotspots in rheumatoid arthritis

### 2024-03-07 scomatic/reports/targeted_mutation_calling.Rmd
add nextflow wrapper to targeted_mutation_calling by id/celltype

```{r mut_hotspots}
# read in bam mappings, get celltype-split bams only
mappings <-
  readr::read_csv(params$bam_mappings) %>%
  dplyr::filter(!is.na(celltype)) %>%
  # temporarily remove problematic BRI-401 endo bam
  dplyr::filter(!(id == 'BRI_401' & celltype == 'Endothelial_cell'))

```{r}
# read targeted nanoseq 
tnanoseq <-
  '/lustre/scratch125/casm/team268im/lh22/synovium_nanoseq/ex_t_dndsout/dnds_nano_muts.tsv' %>%
  readr::read_tsv()

# read genes of interest
target_genes <-
  '/lustre/scratch125/casm/team268im/at31/RA_som_mut/scomatic/data/driver_genes/Sanger_TERT-v4_TE-95148282_hg19_highstringencyfilter_buccal_gene_list.tsv' %>%
  readr::read_tsv()

# run dndscv
dnds <-
  dndscv::dndscv(
    unique(tnanoseq[, 1:5]),
    max_muts_per_gene_per_sample = Inf,
    max_coding_muts_per_sample = Inf,
    gene_list = target_genes$gene,
    outmats = T)

@@

          recursites %>%
            # dplyr::filter(gene == 'TNFRSF14', aachange == 'W12*') %>% 
            purrr::pmap(function(chr, pos, start20, stop20, gene, ...) {
              # testing: # recursites %>% dplyr::filter(gene == 'TNFRSF14', aachange == 'W12*', pos == mut_pos) %>% as.list() %>% list2env(envir = globalenv())

@@

              # get reads at each position of the window
              deepSNV::bam2R(
                ct_bam,
                chr = chr,
                # extend pileup window 20bp either side of mutated position
                start = pos - 20,
                stop = pos + 20,
                q = min_phred,
                mask = sam_flag,
                mq = mapq) %>%
                tibble::as_tibble() %>%
                dplyr::mutate(chr = chr, pos = positions, mut_pos = pos, gene = gene) %>%
                dplyr::select(chr, pos, mut_pos, gene, dplyr::all_of(nucleotides))

@@

>>>>

```{r load}
# read targeted nanoseq 
tnanoseq <-
  '/lustre/scratch125/casm/team268im/lh22/synovium_nanoseq/ex_t_dndsout/dnds_nano_muts.tsv' %>%
  readr::read_tsv()

# read genes of interest
target_genes <-
  '/lustre/scratch125/casm/team268im/at31/RA_som_mut/scomatic/data/driver_genes/Sanger_TERT-v4_TE-95148282_hg19_highstringencyfilter_buccal_gene_list.tsv' %>%
  readr::read_tsv()
```

```{r dndscv_and_liftover}
# run dndscv
dnds <-
  dndscv::dndscv(
    unique(tnanoseq[, 1:5]),
    max_muts_per_gene_per_sample = Inf,
    max_coding_muts_per_sample = Inf,
    gene_list = target_genes$gene,
    outmats = T)

@@

          recursites %>%
            # dplyr::filter(gene == 'TNFRSF14', aachange == 'W12*') %>% 
            purrr::pmap(function(chr, pos, ref, mut, gene, ...) {
              # testing: # recursites %>% dplyr::filter(gene == 'TNFRSF14', aachange == 'W12*') %>% as.list() %>% list2env(envir = globalenv())

@@

              # get reads at each position of the window
              ct_bam %>%
                deepSNV::bam2R(
                  chr = chr,
                  # extend pileup window 20bp either side of mutated position
                  start = pos - 20,
                  stop = pos + 20,
                  # quality controls
                  q = min_phred,
                  mask = sam_flag,
                  mq = mapq) %>%
                tibble::as_tibble() %>%
                dplyr::mutate(chr = chr, pos = positions, mut_pos = pos, gene = gene) %>%
                dplyr::select(chr, pos, mut_pos, gene, dplyr::all_of(nucleotides))

@@

```{r mut_hotspots, eval = F}
# read in bam mappings, get celltype-split bams only
mappings <-
  readr::read_csv(params$bam_mappings) %>%
  dplyr::filter(!is.na(celltype)) %>%
  # temporarily remove problematic BRI-401 endo bam
  dplyr::filter(!(id == 'BRI_401' & celltype == 'Endothelial_cell'))
